# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r loading}
library(knitr)
library(ggplot2)
library(reshape2)

setwd("C:/Users/Petya/Dropbox/coursera/ReproducibleResearch/PeerAssessment1/raw")
activity <- read.csv("../data/activity.csv")
str(activity)
activity$interval <- factor(activity$interval)
activity$date <- as.Date(activity$date, format="%Y-%m-%d")
```

## What is mean total number of steps taken per day?
```{r hist}
hist(activity$steps, main="Histogram of steps taken each day", xlab="Number of steps")
steps_mean <- mean(activity$steps, na.rm=TRUE)
steps_median <- median(activity$steps, na.rm=TRUE)
```
The mean total number of steps taken per day is `r steps_mean`, and the median is `r steps_median`.


## What is the average daily activity pattern?
```{r pattern}
adap <- as.data.frame(tapply(activity$steps, activity$interval, mean, na.rm=TRUE))
names(adap) <- "adap"
plot(x=row.names(adap), y=adap$adap,type = "l", xlab="Time interval", ylab="Average number of steps")
adap_max <- row.names(adap)[which(adap$adap==max(adap$adap))]
```

The 5-minute interval that contains the maximum number of steps, on average across all the days in the dataset is `r adap_max`.

## Imputing missing values
```{r NAs}
num_na <- sum(is.na(activity$steps))
activity2 <- activity
activity2$steps[is.na(activity2$steps)] <- adap$adap[match(activity2$interval[is.na(activity2$steps)],row.names(adap))]

hist(activity2$steps, main="Histogram of steps taken each day", xlab="Number of steps")
steps_mean2 <- mean(activity2$steps)
steps_median2 <- median(activity2$steps)
```

The number of missing values in the steps column is `r num_na`.
We imputed missing values with the daily average of the corresponding interval.
The new mean is `r steps_mean2` and the new median is `r steps_median2`. Both are the same as the estimates without the missing values, so imputation had no effect on these estimates.


## Are there differences in activity patterns between weekdays and weekends?
```{r weekday}
activity2$day <- factor(weekdays(activity2$date) %in% c("szombat","vasárnap"), labels=c("weekday","weekend")) #szombat is saturday, vasárnap is sunday

adap2 <- as.data.frame(tapply(activity2$steps, list(activity2$interval, activity2$day), mean, na.rm=TRUE))
adap2$interval <- rownames(adap2)
adap2_long <- melt(adap2)
adap2_long$interval <- as.numeric(adap2_long$interval)

ggplot(data=adap2_long, aes(x=interval,y=value))+geom_line()+facet_grid(variable~.)+ylab("Number of steps")
```
